\boldmath Disconnected pseudo-$C_\ell$ covariances for projected large-scale structure data

### 2019-09-26 draft/preprint_JCAP_v0.2/paper.tex
first inclusion of revisor comments to v0.1

\newcommand{\nv}{\hat{\boldsymbol{\theta}}}
\newcommand{\summ}[1]{\sum_{\bf #1}\Delta #1^2}
\newcommand{\Ylm}[3]{{\sf E}^{#1}_{\bf #2}({\bf #3})}
\newcommand{\wtj}[6]{\left(\begin{array}{ccc} #1 & #2 & #3\\#4 & #5 & #6\end{array} \right)}

@@

\abstract{The disconnected part of the power spectrum covariance matrix (also known as the ``Gaussian'' covariance) is the dominant contribution on large scales for galaxy clustering and weak lensing datasets. The presence of a complicated sky mask causes non-trivial correlations between different Fourier/harmonic modes, which must be accurately characterized in order to obtain reliable cosmological constraints. This is particularly relevant for galaxy survey data. Unfortunately, an exact calculation of these correlations involves $O(\ell_{\rm max}^6)$ operations that become computationally impractical very quickly. We present an implementation of approximate methods to estimate the Gaussian covariance matrix of power spectra involving spin-0 and spin-2 flat- and curved-sky fields, expanding on existing algorithms. These methods achieve an $O(\ell_{\rm max}^3)$ scaling, which makes the computation of the covariance matrix as fast as the computation of the power spectrum itself. We quantify the accuracy of these methods on large-scale structure and weak lensing data, making use of a large number of Gaussian but otherwise realistic simulations. We show that, using the approximate covariance matrix, we are able to recover the true posterior distribution of cosmological parameters to high accuracy. We also quantify the shortcomings of these methods, which become unreliable on the very largest scales, as well as for covariance matrix elements involving cosmic shear $B$ modes. The algorithms presented here are implemented in the public code {\tt NaMaster} \url{https://github.com/LSSTDESC/NaMaster}.}

@@

  \section{Introduction}\label{sec:intro}
    The two-point correlation of different fields projected on the celestial sphere is one of the most common observables used in the analysis of large datasets in astrophysics, from studies of the Cosmic Microwave Background \citep{1995PhRvL..74.4369B,1997PhRvD..55.5895T,1998PhRvD..57.2117B,2001PhRvD..64f3001T,2002ApJ...567....2H,2003PhRvD..67b3001W} to large-scale structure and weak lensing surveys \cite{2000MNRAS.317L..23H,2001ApJ...555..547H,2001MNRAS.325.1603E,2011arXiv1112.5723H,2016MNRAS.456.1508K,2018MNRAS.476.1050B,2018arXiv181208182X,2019MNRAS.tmp.1446C,2019PASJ...71...43H}. Using these two-point functions, one achieves a high level of data compression (with respect to the size of the raw datasets -- time-ordered data, images or catalogs). They can also be directly used to constrain cosmological and astrophysical parameters assuming that one can model their likelihood. This is usually done by assuming that the two-point functions are Gaussianly distributed, which is often a good approximation due to the central limit theorem \cite{2008PhRvD..77j3013H,2018MNRAS.477.4879S}. In this case, the only obstacle that remains is being able to estimate the covariance matrix of a set of two-point correlators. Since the form of this covariance directly affects the posterior parameter uncertainties, a precise determination of it is of paramount importance. In large-scale structure experiments, this has often been resolved by making use of one's own data through resampling techniques \cite{tukey1958,1984MNRAS.210P..19B,1993ApJ...406L..47H,2002ApJ...571..172Z,2016arXiv160600233E}, or by generating a large number of mock realizations \cite{2013MNRAS.428.1036M,2018MNRAS.479...94A,2019MNRAS.485.2806B}. With the advent of the larger current and future surveys, the increasing size of the data vector and of the volume to be simulated has made this solution impractical, and fully analytical and hybrid estimators are now being used.

@@

      Now, in what follows, we will be concerned with the auto- and
      cross-correlations of spin-0 and spin-2 fields. Thus, to simplify the
      notation, we will enumerate the different types of coupling coefficients
      that exist for a spin-0 field with a single component that we will call
      $\delta$, in analogy to the projected galaxy overensity, and for a
      spin-2 field, $\boldsymbol{\gamma}$, with $E$ and $B$ components, $\gamma_E$ and $\gamma_B$, in analogy to the cosmic shear field. With this setup, all the possible non-zero $W^{\alpha\beta,\alpha'\beta'}_{{\bf l}{\bf l}'}$ can be expressed in terms of the $I^{\pm s_a,\pm s_b}_{{\bf l}{\bf l}'}$ as follows:
      \begin{align}\label{eq:coeff}
        &W^{\delta\delta,\delta\delta}_{{\bf l},{\bf l}'}=I^{0,0}_{{\bf l}{\bf l}'}; \\
        &W^{\delta\gamma_E,\delta\gamma_E}_{{\bf l},{\bf l}'}=W^{\delta\gamma_B,\delta\gamma_B}_{{\bf l},{\bf l}'}=I^{0,+2}_{{\bf l},{\bf l}'};\hspace{12pt}
        W^{\delta\gamma_E,\delta\gamma_B}_{{\bf l},{\bf l}'}=-W^{\delta\gamma_B,\delta\gamma_E}_{{\bf l},{\bf l}'}=-i\,I^{0,-2}_{{\bf l},{\bf l}'};\\
        &W^{\gamma_E\gamma_E,\gamma_E\gamma_E}_{{\bf l},{\bf l}'}=W^{\gamma_E\gamma_B,\gamma_E\gamma_B}_{{\bf l},{\bf l}'}=W^{\gamma_B\gamma_E,\gamma_B\gamma_E}_{{\bf l},{\bf l}'}=W^{\gamma_B\gamma_B,\gamma_B\gamma_B}_{{\bf l},{\bf l}'}=I^{+2,+2}_{{\bf l}{\bf l}'}\\
        &W^{\gamma_E\gamma_E,\gamma_B\gamma_B}_{{\bf l},{\bf l}'}=W^{\gamma_B\gamma_B,\gamma_E\gamma_E}_{{\bf l},{\bf l}'}=-W^{\gamma_B\gamma_E,\gamma_E\gamma_B}_{{\bf l},{\bf l}'}=-W^{\gamma_E\gamma_B,\gamma_B\gamma_E}_{{\bf l},{\bf l}'}=I^{-2,-2}_{{\bf l}{\bf l}'}\\
        &W^{\gamma_E\gamma_E,\gamma_E\gamma_B}_{{\bf l},{\bf l}'}=-W^{\gamma_E\gamma_B,\gamma_E\gamma_E}_{{\bf l},{\bf l}'}=W^{\gamma_B\gamma_E,\gamma_B\gamma_B}_{{\bf l},{\bf l}'}=-W^{\gamma_B\gamma_B,\gamma_B\gamma_E}_{{\bf l},{\bf l}'}=-i\,I^{+2,-2}_{{\bf l}{\bf l}'}\\
        &W^{\gamma_E\gamma_E,\gamma_B\gamma_E}_{{\bf l},{\bf l}'}=W^{\gamma_E\gamma_B,\gamma_B\gamma_B}_{{\bf l},{\bf l}'}=-W^{\gamma_B\gamma_E,\gamma_E\gamma_E}_{{\bf l},{\bf l}'}=-W^{\gamma_B\gamma_B,\gamma_E\gamma_B}_{{\bf l},{\bf l}'}=i\,I^{-2,+2}_{{\bf l}{\bf l}'}.
      \end{align}
      Thus, in principle, we only need to compute 7 different types of terms
      ($I^{0,0}$, $I^{0,\pm2}$, $I^{\pm2,\pm2}$ and $I^{\pm2,\mp2})$). In
      order to simplify these expressions further, we follow \cite{2017A&A...602A..41C} and neglect all gradients of the masks, which allows us to relate the different $I^{\pm s_a,\pm s_b}$ through the following set of identities:
      \begin{align}\nonumber
        \summ{x}w({\bf x})\left(\bar{\eth}^2q({\bf l},{\bf x})\right)^*\,\bar{\eth}^2q({\bf k},{\bf x})
        &=\summ{x}w({\bf x})\eth^2 q^*({\bf l},{\bf x})\,\bar{\eth}^2q({\bf k},{\bf x})\\\nonumber
        &=\summ{x}q^*({\bf l},{\bf x})\,\eth^2\left(\bar{\eth}^2q({\bf k},{\bf x})w({\bf x})\right)\\\nonumber
        &\simeq\summ{x}q^*({\bf l},{\bf x})\,\left(\eth^2\bar{\eth}^2q({\bf k},{\bf x})\right)w({\bf x})\\
        &=\frac{1}{\beta^2_{k,2}}\summ{x}q^*({\bf l},{\bf x})\,q({\bf k},{\bf x})w({\bf x})\\\nonumber
        &=\summ{x}q^*({\bf l},{\bf x})\,\left(\bar{\eth}^2\eth^2q({\bf k},{\bf x})\right)w({\bf x})\\\nonumber
        &=\summ{x}\bar{\eth}^2\left(w({\bf x})\,q^*({\bf l},{\bf x})\right)\,\eth^2q({\bf k},{\bf x})\\
        &\simeq\summ{x}w({\bf x})\,\left(\eth^2\,q({\bf l},{\bf x})\right)^*\,\eth^2q({\bf k},{\bf x}),
      \end{align}
      where we have made repeated use of integration by parts.

@@

      For our fiducial results, we generated a set of 20,000 random simulations. This number was chosen in order to recover the covariance matrix for all possible auto- and cross-correlations with sufficient accuracy. All simulations were generated as {\tt HEALPix} maps with resolution $N_{\rm side}=512$. We then used {\tt NaMaster} to compute all possible power spectra for each simulation using narrow bandpowers of width $\Delta\ell=3$ from $\ell=2$ to $\ell=1023$. Finally, we used the power spectra from the simulations to estimate the sample covariance matrix
      \begin{equation}
        {\rm Cov}({\bf C}) =\frac{1}{N_{\rm sim}-1}\sum_{i=1}^{N_{\rm sim}}\left({\bf C}_i-\bar{\bf C}\right)\cdot\left({\bf C}_i-\bar{\bf C}\right)^T,
      \end{equation}
      where ${\bf C}_i$ is the vector of all possible power spectra for the $i$-th realization, and $\bar{\bf C}$ is the mean of this vector over all realizations. The comparison of this sample covariance with the analytical approximations described above is presented in the next sections.

@@

      \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{./figures/contaminants_diag_cov.pdf}
        \caption{Diagonal of the covariance matrix estimated from simulations without contaminants (solid blue), with 200 contaminants deprojected (dashed orange) and the NKA estimator (solid green). The differences between the simulated cases are significantly smaller than those associated with the NKA estimator. The NKA method is therefore equally valid to approximate the covariance even in the presence of contaminant deprojection.}
        \label{fig:diag_conts}
      \end{figure}
      Figure \ref{fig:diag_conts} serves also to illustrate another important point. In all our tests we find that the largest differences between the sample and NKA covariances occur on large scales. For $\delta$-$\delta$ correlations, the effect is limited to the first few multipoles ($\ell\lesssim10$), while for $\gamma_E$-$\gamma_E$ we are only able to recover the sample covariance errors within $5\%$ for $\ell\gtrsim40$. This is the main source of the small mismatch observed in Figure~\ref{fig:chi2_1bin}. Note, however, that most of the cosmological information is obtained from the higher multipoles, due to their higher statistical weight, and, as we will show in the next section, the effect on the final parameter constraints is negligible. If accurate covariances bars are needed on these large scales for cosmic shear, they can be estimated alternatively making use of fast, low resolution simulations (e.g. {\tt HEALPix} $N_{\rm side}=64$ maps).

    \subsection{Impact on parameter estimation}\label{ssec:results.cosmo}
      \begin{figure}
        \centering
        \includegraphics[width=0.6\textwidth]{./figures/contours_2bin.pdf}
        \caption{1$\sigma$ and 2$\sigma$ contours for $\Omega_M$ and $\sigma_8$ found from two different estimates of the power spectrum likelihood for a dataset containing all possible cross-power spectra between $\delta$ and $\gamma_E$ for two redshift bins. The filled contours correspond to the parameter likelihood evaluated using the NKA estimator for the power spectrum covariance matrix, while the dashed black contours correspond to the same calculation done with the sample covariance matrix corrected by the factor in Eq. \ref{eq:hartlap}. In both cases we assume a Gaussian likelihood. We find a remarkably good agreement between both likelihoods, highlighting the negligible impact of the approximations involved in the NKA method on the final parameter estimates.} \label{fig:contours}
      \end{figure}
      Ultimately, the most important test to judge the accuracy of the analytical covariance matrix estimators implemented here is to study their impact on the posterior distribution of cosmological parameters derived from power spectrum measurements. Assuming flat priors, the posterior distribution for parameters $\vec{\theta}$ is simply given by:
      \begin{equation}\label{eq:parpos}
        -2\log p(\vec{\theta}|\hat{\bf C})=\left(\hat{\bf C}-{\bf C}(\vec{\theta})\right)^T\cdot{\rm Cov}^{-1}\cdot\left(\hat{\bf C}-{\bf C}(\vec{\theta})\right)+{\rm const.},
      \end{equation}
      where $\hat{\bf C}$ is a vector of power spectrum measurements, ${\rm Cov}$ is their covariance matrix, and ${\bf C}(\vec{\theta})$ is their theoretical prediction for parameters $\vec{\theta}$.

      We explore $\log p(\vec{\theta}|\hat{\bf C})$ for of two parameters $\vec{\theta}\equiv(\Omega_M,\sigma_8)$, for a data vector composed of all possible auto- and cross-correlations between $\delta$ and $\gamma_E$ in the case of two redshift bins described in Section \ref{ssec:results.sims}. In this simple two-dimensional scenario, we simply sample the distribution in a regular grid of 100 by 100 points for each parameter. We construct a data vector $\hat{\bf C}$ from the theoretical prediction for the experimental setup described in Section \ref{ssec:results.sims}, and produce theoretical predictions for it at each grid point using {\tt CCL}. Note that, when evaluating the posterior for the sample covariance matrix, one needs to correct for the finite number of simulations used to construct the covariance. In most situations this can be done simply by rescaling the inverse covariance matrix by a factor given by \cite{2007A&A...464..399H}
      \begin{equation}\label{eq:hartlap}
        {\rm Cov}^{-1}\,\longrightarrow \frac{N_s-2-N_{\rm data}}{N_s-1}\,{\rm Cov}^{-1},
      \end{equation}
      where $N_s=20,000$ is the number of samples used to estimate the covariance, and $N_{\rm data}=3510$ is the number of data points.

@@

    More in detail, we have also found that the impact of contaminant deprojection on the covariance matrix, through the corresponding loss of modes, is negligible unless a very large number of contaminant templates are removed. This simplifies the procedure to estimate covariance matrices for galaxy clustering data, which are particularly sensitive to a large number of astrophysical and observational systematics. Additionally, we have found that, although the NKA estimator is accurate enough, it is not able to perfectly capture the additional effects of mode coupling that are present for spin-2 fields, and that a simpler approach treating the shear $E$ modes as a spin-0 object is able to reach similar levels of accuracy. Due to the imperfect treatment of the $E/B$ mixing caused by the sky mask in the NKA estimator, we also find that the predicted covariance matrix for any power spectra involving $B$ modes differs significantly from the true sample covariance, and, therefore, this approach cannot be used to reliably estimate the uncertainties of $B$-mode power spectra. Likewise, we find that the NKA estimator yields inaccurate estimates of the power spectrum uncertainties on the largest scales ($\ell\lesssim50$). Although these modes carry a substantially smaller statistical weight, if more accurate covariances are needed on these large scales, they can be easily computed making use of fast low-resolution simulations.

@@

  \acknowledgments
  We thank Thibaut Louis, Eva-Maria Mueller, Francisco Javier S\'anchez and An\v ze Slosar for useful comments and discussion. CGG is supported the Spanish grant BES-2016-077038, partially funded by the ESF and by AYA2015-67854-P from the Ministry of Industry, Science and Innovation of Spain and the FEDER funds. He was partially supported by a Balzan Fellowship while in Oxford. He would like to thank New College and the Department of Physics at Oxford for their hospitality. DA acknowledges support from STFC through an Ernest Rutherford Fellowship, grant reference ST/P004474/1. EB acknowledges support from the Beecroft Trust.

>>>>

\newcommand{\cgg}[1]{{\color{olive} #1}}

\newcommand{\nv}{\hat{\boldsymbol{\theta}}}
\newcommand{\summ}[1]{\sum_{\bf #1}\Delta^2 #1\,}
\newcommand{\Ylm}[3]{{\sf E}^{#1}_{\bf #2}({\bf #3})}
\newcommand{\wtj}[6]{\left(\begin{array}{ccc} #1 & #2 & #3\\#4 & #5 & #6\end{array} \right)}

@@

\abstract{The disconnected part of the power spectrum covariance matrix (also
  known as the ``Gaussian'' covariance) is the dominant contribution on large
  scales for galaxy clustering and weak lensing datasets. The presence of a
  complicated sky mask causes non-trivial correlations between different
  Fourier/harmonic modes, which must be accurately characterized in order to
  obtain reliable cosmological constraints. This is particularly relevant for
  galaxy survey data. Unfortunately, an exact calculation of these
  correlations involves $O(\ell_{\rm max}^6)$ operations that become
  computationally impractical very quickly. We present an implementation of
  approximate methods to estimate the Gaussian covariance matrix of power
  spectra involving spin-0 and spin-2 flat- and curved-sky fields, expanding
  on existing algorithms \cgg{for CMB}. These methods achieve an $O(\ell_{\rm max}^3)$
  scaling, which makes the computation of the covariance matrix as fast as the
  computation of the power spectrum itself. We quantify the accuracy of these
  methods on large-scale structure and weak lensing data, making use of a
  large number of Gaussian but otherwise realistic simulations. We show that,
  using the approximate covariance matrix, we are able to recover the true
  posterior distribution of cosmological parameters to high accuracy. We also
  quantify the shortcomings of these methods, which become unreliable on the
  very largest scales, as well as for covariance matrix elements involving
  cosmic shear $B$ modes. The algorithms presented here are implemented in the
  public code {\tt NaMaster} \url{https://github.com/LSSTDESC/NaMaster}.}

@@

  \section{Introduction}\label{sec:intro}
    The two-point correlation of different fields projected on the celestial
    sphere is one of the most common observables used in the analysis of large
    datasets in astrophysics, from studies of the Cosmic Microwave Background
    \cite{Gorski:1994ye,Gorski:1994uu,1995PhRvL..74.4369B,Gorski:1996cf,Gorski:1996ti,1997PhRvD..55.5895T,1998PhRvD..57.2117B,Wandelt:2000av,2001PhRvD..64f3001T,2002ApJ...567....2H,2003PhRvD..67b3001W}
    to large-scale structure and weak lensing surveys
    \cite{2000MNRAS.317L..23H,2001ApJ...555..547H,2001MNRAS.325.1603E,2011arXiv1112.5723H,2016MNRAS.456.1508K,2018MNRAS.476.1050B,2018arXiv181208182X,2019MNRAS.tmp.1446C,2019PASJ...71...43H}.
    Using these two-point functions, one achieves a high level of data
    compression (with respect to the size of the raw datasets -- time-ordered
    data, images or catalogs). They can also be directly used to constrain
    cosmological and astrophysical parameters assuming that one can model
    their likelihood. This is usually done by assuming that the two-point
    functions are Gaussianly distributed, which is often a good approximation
    due to the central limit theorem
    \cite{2008PhRvD..77j3013H,2018MNRAS.477.4879S}. In this case, the only
    obstacle that remains is being able to estimate the covariance matrix of a
    set of two-point correlators. Since the form of this covariance directly
    affects the posterior parameter uncertainties, a precise determination of
    it is of paramount importance. In large-scale structure experiments, this
    has often been resolved by making use of one's own data through resampling
    techniques
    \cite{tukey1958,1984MNRAS.210P..19B,1993ApJ...406L..47H,2002ApJ...571..172Z,2016arXiv160600233E},
    or by generating a large number of mock realizations
    \cite{2013MNRAS.428.1036M,2018MNRAS.479...94A,2019MNRAS.485.2806B}. With
    the advent of the larger current and future surveys, the increasing size
    of the data vector and of the volume to be simulated has made this
    solution impractical, and fully analytical and hybrid estimators are now
    being used.

@@

      Now, in what follows, we will be concerned with the auto- and
      cross-correlations of spin-0 and spin-2 fields. Thus, to simplify the
      notation, we will enumerate the different types of coupling coefficients
      that exist for a spin-0 field with a single component that we will call
      $\delta$, in analogy to the projected galaxy overensity, and for a
      spin-2 field, $\boldsymbol{\gamma}$, with $E$ and $B$ components, $\gamma_E$ and $\gamma_B$, in analogy to the cosmic shear field. With this setup, all the possible non-zero $W^{\alpha\beta,\alpha'\beta'}_{{\bf l}{\bf l}'}$ can be expressed in terms of the $I^{\pm s_a,\pm s_b}_{{\bf l}{\bf l}'}$ as follows:
      \begin{align}\label{eq:coeff}
        &W^{\delta\delta,\delta\delta}_{{\bf l},{\bf l}'}=I^{0,0}_{{\bf l}{\bf l}'}; \\
        &W^{\delta\gamma_E,\delta\gamma_E}_{{\bf l},{\bf l}'}=W^{\delta\gamma_B,\delta\gamma_B}_{{\bf l},{\bf l}'}=I^{0,+2}_{{\bf l},{\bf l}'};\hspace{12pt}
        W^{\delta\gamma_E,\delta\gamma_B}_{{\bf l},{\bf l}'}=-W^{\delta\gamma_B,\delta\gamma_E}_{{\bf l},{\bf l}'}=-i\,I^{0,-2}_{{\bf l},{\bf l}'};\\
        &W^{\gamma_E\gamma_E,\gamma_E\gamma_E}_{{\bf l},{\bf l}'}=W^{\gamma_E\gamma_B,\gamma_E\gamma_B}_{{\bf l},{\bf l}'}=W^{\gamma_B\gamma_E,\gamma_B\gamma_E}_{{\bf l},{\bf l}'}=W^{\gamma_B\gamma_B,\gamma_B\gamma_B}_{{\bf l},{\bf l}'}=I^{+2,+2}_{{\bf l}{\bf l}'}\\
        &W^{\gamma_E\gamma_E,\gamma_B\gamma_B}_{{\bf l},{\bf l}'}=W^{\gamma_B\gamma_B,\gamma_E\gamma_E}_{{\bf l},{\bf l}'}=-W^{\gamma_B\gamma_E,\gamma_E\gamma_B}_{{\bf l},{\bf l}'}=-W^{\gamma_E\gamma_B,\gamma_B\gamma_E}_{{\bf l},{\bf l}'}=I^{-2,-2}_{{\bf l}{\bf l}'}\\
        &W^{\gamma_E\gamma_E,\gamma_E\gamma_B}_{{\bf l},{\bf l}'}=-W^{\gamma_E\gamma_B,\gamma_E\gamma_E}_{{\bf l},{\bf l}'}=W^{\gamma_B\gamma_E,\gamma_B\gamma_B}_{{\bf l},{\bf l}'}=-W^{\gamma_B\gamma_B,\gamma_B\gamma_E}_{{\bf l},{\bf l}'}=-i\,I^{+2,-2}_{{\bf l}{\bf l}'}\\
        &W^{\gamma_E\gamma_E,\gamma_B\gamma_E}_{{\bf l},{\bf l}'}=W^{\gamma_E\gamma_B,\gamma_B\gamma_B}_{{\bf l},{\bf l}'}=-W^{\gamma_B\gamma_E,\gamma_E\gamma_E}_{{\bf l},{\bf l}'}=-W^{\gamma_B\gamma_B,\gamma_E\gamma_B}_{{\bf l},{\bf l}'}=i\,I^{-2,+2}_{{\bf l}{\bf l}'}.
      \end{align}
      Thus, in principle, we only need to compute 7 different types of terms
      ($I^{0,0}$, $I^{0,\pm2}$, $I^{\pm2,\pm2}$ and $I^{\pm2,\mp2})$). In
      order to simplify these expressions further, we follow \cite{Efstathiou:2006eb,2017A&A...602A..41C} and neglect all gradients of the masks, which allows us to relate the different $I^{\pm s_a,\pm s_b}$ through the following set of identities:
      \begin{align}\nonumber
        \summ{x}w({\bf x})\left(\bar{\eth}^2q({\bf l},{\bf x})\right)^*\,\bar{\eth}^2q({\bf k},{\bf x})
        &=\summ{x}w({\bf x})\eth^2 q^*({\bf l},{\bf x})\,\bar{\eth}^2q({\bf k},{\bf x})\\\nonumber
        &=\summ{x}q^*({\bf l},{\bf x})\,\eth^2\left(\bar{\eth}^2q({\bf k},{\bf x})w({\bf x})\right)\\\nonumber
        &\simeq\summ{x}q^*({\bf l},{\bf x})\,\left(\eth^2\bar{\eth}^2q({\bf k},{\bf x})\right)w({\bf x})\\
        &=\frac{1}{\beta^2_{k,2}}\summ{x}q^*({\bf l},{\bf x})\,q({\bf k},{\bf x})w({\bf x})\\\nonumber
        &=\summ{x}q^*({\bf l},{\bf x})\,\left(\bar{\eth}^2\eth^2q({\bf k},{\bf x})\right)w({\bf x})\\\nonumber
        &=\summ{x}\bar{\eth}^2\left(w({\bf x})\,q^*({\bf l},{\bf x})\right)\,\eth^2q({\bf k},{\bf x})\\
        &\simeq\summ{x}w({\bf x})\,\left(\eth^2\,q({\bf l},{\bf x})\right)^*\,\eth^2q({\bf k},{\bf x}),
      \end{align}
      where we have made repeated use of integration by parts.

@@

      For our fiducial results, we generated a set of 20,000 random
      simulations. This number was chosen in order to recover the covariance
      matrix for all possible auto- and cross-correlations with sufficient
      accuracy. All simulations were generated as {\tt
        HEALPix}\footnote{\url{http://healpix.sourceforge.net}} maps with resolution $N_{\rm side}=512$. We then used {\tt NaMaster} to compute all possible power spectra for each simulation using narrow bandpowers of width $\Delta\ell=3$ from $\ell=2$ to $\ell=1023$. Finally, we used the power spectra from the simulations to estimate the sample covariance matrix
      \begin{equation}
        {\rm Cov}({\bf C}) =\frac{1}{N_{\rm sim}-1}\sum_{i=1}^{N_{\rm sim}}\left({\bf C}_i-\bar{\bf C}\right)\cdot\left({\bf C}_i-\bar{\bf C}\right)^T,
      \end{equation}
      where ${\bf C}_i$ is the vector of all possible power spectra for the $i$-th realization, and $\bar{\bf C}$ is the mean of this vector over all realizations. The comparison of this sample covariance with the analytical approximations described above is presented in the next sections.

@@

      \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{./figures/contaminants_diag_cov.pdf}
        \caption{Diagonal of the covariance matrix estimated from simulations without contaminants (solid blue), with 200 contaminants deprojected (dashed orange) and the NKA estimator (solid green). The differences between the simulated cases are significantly smaller than those associated with the NKA estimator. The NKA method is therefore equally valid to approximate the covariance even in the presence of contaminant deprojection.}
        \label{fig:diag_conts}
      \end{figure}
      Figure \ref{fig:diag_conts} serves also to illustrate another important
      point. In all our tests we find that the largest differences between the
      sample and NKA covariances occur on large scales. For $\delta$-$\delta$
      correlations, the effect is limited to the first few multipoles
      ($\ell\lesssim10$), while for $\gamma_E$-$\gamma_E$ we are only able to
      recover the sample covariance errors within $5\%$ for $\ell\gtrsim40$.
      This is the main source of the small mismatch observed in
      Figure~\ref{fig:chi2_1bin}. Note, however, that most of the cosmological
      information is obtained from the higher multipoles, due to their higher
      statistical weight, and, as we will show in the next section, the effect
      on the final parameter constraints is negligible. If accurate
      \cgg{covariance matrix elements} are needed on these large scales for cosmic shear, they
      can be estimated alternatively making use of fast, low resolution
      simulations (e.g. {\tt HEALPix} $N_{\rm side}=64$ maps) \cgg{or computed
      exactly as in Ref.~\cite{Efstathiou:2006eb,2017A&A...602A..41C}}.

    \subsection{Impact on parameter estimation}\label{ssec:results.cosmo}
      \begin{figure}
        \centering
        \includegraphics[width=0.6\textwidth]{./figures/contours_2bin.pdf}
        \caption{1$\sigma$ and 2$\sigma$ contours for $\Omega_M$ and $\sigma_8$ found from two different estimates of the power spectrum likelihood for a dataset containing all possible cross-power spectra between $\delta$ and $\gamma_E$ for two redshift bins. The filled contours correspond to the parameter likelihood evaluated using the NKA estimator for the power spectrum covariance matrix, while the dashed black contours correspond to the same calculation done with the sample covariance matrix corrected by the factor in Eq. \ref{eq:hartlap}. In both cases we assume a Gaussian likelihood. We find a remarkably good agreement between both likelihoods, highlighting the negligible impact of the approximations involved in the NKA method on the final parameter estimates.} \label{fig:contours}
      \end{figure}
      Ultimately, the most important test to judge the accuracy of the
      analytical covariance matrix estimators implemented here is to study
      their impact on the posterior distribution of cosmological parameters
      derived from power spectrum measurements. Assuming flat priors \cgg{and
      a fiducial Gaussian likelihood approximation~\cite{2008PhRvD..77j3013H}}, the posterior distribution for parameters $\vec{\theta}$ is simply given by:
      \begin{equation}\label{eq:parpos}
        -2\log p(\vec{\theta}|\hat{\bf C})=\left(\hat{\bf C}-{\bf C}(\vec{\theta})\right)^T\cdot{\rm Cov}^{-1}\cdot\left(\hat{\bf C}-{\bf C}(\vec{\theta})\right)+{\rm const.},
      \end{equation}
      where $\hat{\bf C}$ is a vector of power spectrum measurements, ${\rm Cov}$ is their covariance matrix, and ${\bf C}(\vec{\theta})$ is their theoretical prediction for parameters $\vec{\theta}$.

      We explore $\log p(\vec{\theta}|\hat{\bf C})$ for \cgg{the} two parameters $\vec{\theta}\equiv(\Omega_M,\sigma_8)$, for a data vector composed of all possible auto- and cross-correlations between $\delta$ and $\gamma_E$ in the case of two redshift bins described in Section \ref{ssec:results.sims}. In this simple two-dimensional scenario, we simply sample the distribution in a regular grid of 100 by 100 points for each parameter. We construct a data vector $\hat{\bf C}$ from the theoretical prediction for the experimental setup described in Section \ref{ssec:results.sims}, and produce theoretical predictions for it at each grid point using {\tt CCL}. Note that, when evaluating the posterior for the sample covariance matrix, one needs to correct for the finite number of simulations used to construct the covariance. In most situations this can be done simply by rescaling the inverse covariance matrix by a factor given by \cite{2007A&A...464..399H}
      \begin{equation}\label{eq:hartlap}
        {\rm Cov}^{-1}\,\longrightarrow \frac{N_s-2-N_{\rm data}}{N_s-1}\,{\rm Cov}^{-1},
      \end{equation}
      where $N_s=20,000$ is the number of samples used to estimate the covariance, and $N_{\rm data}=3510$ is the number of data points.

@@

    More in detail, we have also found that the impact of contaminant
    deprojection on the covariance matrix, through the corresponding loss of
    modes, is negligible unless a very large number of contaminant templates
    are removed. This simplifies the procedure to estimate covariance matrices
    for galaxy clustering data, which are particularly sensitive to a large
    number of astrophysical and observational systematics. Additionally, we
    have found that, although the NKA estimator is accurate enough, it is not
    able to perfectly capture the additional effects of mode coupling that are
    present for spin-2 fields, and that a simpler approach treating the shear
    $E$ modes as a spin-0 object is able to reach similar levels of accuracy.
    Due to the imperfect treatment of the $E/B$ mixing caused by the sky mask
    in the NKA estimator, we also find that the predicted covariance matrix
    for any power spectra involving $B$ modes differs significantly from the
    true sample covariance, and, therefore, this approach cannot be used to
    reliably estimate the uncertainties of $B$-mode power spectra. Likewise,
    we find that the NKA estimator yields inaccurate estimates of the power
    spectrum uncertainties on the largest scales ($\ell\lesssim50$). Although
    these modes carry a substantially smaller statistical weight, if more
    accurate covariances are needed on these large scales, they can be easily
    computed making use of fast low-resolution simulations \cgg{or exactly
    following Refs.~\cite{Efstathiou:2006eb,2017A&A...602A..41C}}.

@@

  \acknowledgments
  We thank Thibaut Louis, Eva-Maria Mueller, Francisco Javier S\'anchez and
  An\v ze Slosar for useful comments and discussion. CGG is supported the
  Spanish grant BES-2016-077038, partially funded by the ESF and by
  AYA2015-67854-P from the Ministry of Industry, Science and Innovation of
  Spain and the FEDER funds. He was partially supported by a Balzan Fellowship
  while in Oxford. He would like to thank New College and the Department of
  Physics at Oxford for their hospitality. DA acknowledges support from STFC
  through an Ernest Rutherford Fellowship, grant reference ST/P004474/1. EB
  acknowledges support from the Beecroft Trust.  \cgg{Some of the results in this
  paper have been derived using the HEALPix~\cite{Gorski:2004by} package}
